maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-04-09 docs/paper/paper.tex
add benchmarks links, change title, trying to fix section autorefs

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={N/A},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
et al. (1970). maze-dataset. \emph{Journal of Open Source Software},
\emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}

@@

\title{maze-dataset}

@@

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
  Wu, 2023}), but provides over 5.8 million unique formats for
  converting mazes to a text stream, detailed in
  \autoref{tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use a highly efficient method detailed in
  \autoref{implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

\hypertarget{tokenized-output-formats}{%
\subsection{Tokenized Output Formats}\label{tokenized-output-formats}}

@@

\begin{figure}
\hypertarget{fig:benchmarks}{%
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
\caption{Plot of maze generation time. Generation time scales
exponentially with maze size for all algorithms. Generation time per
maze does not depend on the number of mazes being generated, and there
is minimal overhead to initializing the generation process for a small
dataset. Wilson's algorithm is notably less efficient than others and
has high variance. Note that values are averaged across all parameter
sets for that algorithm.}\label{fig:benchmarks}
}
\end{figure}

\hypertarget{success-rate-estimation}{%
\subsection{Success Rate Estimation}\label{success-rate-estimation}}

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
\caption{An example of both empirical and predicted success rates as a
function of the percolation probability \(p\) for various maze sizes,
percolation with and without depth first search, and
\texttt{endpoint\_kwargs} requiring that both the start and end be in
unique dead ends.}
\end{figure}

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

@@

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be controlled via
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
and complications caused by this are detailed in
\autoref{success-rate-estimation}.

>>>>

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alex F.
Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={N/A},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alex F. Spies, Tilman Räuker, Brandon Knutson, Cecilia
Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

@@

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
et al. (1970). maze-dataset: Maze Generation with Algorithmic Variety
and Representational Flexibility. \emph{Journal of Open Source
Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE? \url{https://doi.org/N/A}}.}}

@@

\title{maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility}

@@

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
  Wu, 2023}), but provides over 5.8 million unique formats for
  converting mazes to a text stream, detailed in
  \autoref{sec:tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use a highly efficient method detailed in
  \autoref{sec:implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

\hypertarget{sec:tokenized-output-formats}{%
\subsection{Tokenized Output
Formats}\label{sec:tokenized-output-formats}}

@@

\begin{figure}
\hypertarget{fig:benchmarks}{%
\centering
\includegraphics[width=0.95\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
\caption{Plot of maze generation time. Generation time scales
exponentially with maze size for all algorithms. Generation time per
maze does not depend on the number of mazes being generated, and there
is minimal overhead to initializing the generation process for a small
dataset. Wilson's algorithm is notably less efficient than others and
has high variance. Note that values are averaged across all parameter
sets for that algorithm. More information can be found on the
\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
page}.}\label{fig:benchmarks}
}
\end{figure}

\hypertarget{sec:success-rate-estimation}{%
\subsection{Success Rate Estimation}\label{sec:success-rate-estimation}}

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/ep/ep_deadends_unique-crop.pdf}
\caption{An example of both empirical and predicted success rates as a
function of the percolation probability \(p\) for various maze sizes,
percolation with and without depth first search, and
\texttt{endpoint\_kwargs} requiring that both the start and end be in
unique dead ends. Empirical measures derived from a sample of 128 mazes.
More information can be found on the
\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks
page}.}
\end{figure}

\hypertarget{sec:implementation}{%
\section{Implementation}\label{sec:implementation}}

@@

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be controlled via
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
and complications caused by this are detailed in
\autoref{sec:success-rate-estimation}.
